Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-05-13 Nicolas.tex
48h to go

\usepackage[dvipsnames]{xcolor}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
\newcommand{\nnn}[1]{{\textcolor[rgb]{0.40, 0.0, 0.20}{#1}}}
\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}

@@

\author{N. Nicolas
    \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
    \inst{1}
    \and M. Rigault
    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
    \inst{2}\and\\
    Y. Copin\inst{1}   
    \and R. Graziani\inst{2}
    %HERE ALPHABETIC ORDER
    \and G. Aldering\inst{3}
    \and M. Briday\inst{1}
    \and J. Nordin\inst{4}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    % Gerard ne souhaite pas etre co-auteur.
}

@@

\abstract{
The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
survey statistics increase, the question of astrophysical systematic
uncertainties rises, notably that of the SN~Ia population redshift drift. In
this paper, we study the redshift evolution of SN~Ia \textsc{\texttt{SALT2.4}}
lightcurve stretch, a purely intrinsic SN property, to probe this drift. The SN
stretch has been shown to strongly correlate with the SN's environment, notably
with stellar age tracers. Following the evolution of fraction of young and old
SNe~Ia as predicted by \cite{rigault2018}, and assuming constant underlying
stretch distribution for each population as a function of redshift, model the
expected SN~Ia stretch distribution as a function of redshift. We test our
prediction against data from the literature chosen to have negligible selection
effects to ensure that any observed drift is indeed astrophysical and not
observational. We clearly demonstrate that the underlying SNe~Ia stretch
distribution is evolving as a function of redshift, and that the young/old
drifting model is a much better description of the data than any non-drifting
model, including the sample-based asymmetric distributions used by the Beams
with Bias Correction algorithm to correct Malmquist bias. Our favored underlying
stretch model is bimodal: a high-stretch mode shared by both young and old
environments, and a low-stretch mode exclusive to old environments. The precise
impact of the redshift evolution of the underlying SNe Ia population on
cosmology remains to be studied. Yet, the astrophysical drift of the SN stretch
distribution does affect current Malmquist bias corrections and thereby
distances derived from SN affected by selection effects. We highlight that such
a bias will increase with surveys covering larger redshift ranges, which is
particularly important for LSST.
}

@@

%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.

The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). In \cite{rigault2018} we used the LsSFR to classify which are the
younger (those with a high LsSFR) and which are the older (those with low
LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
been known to be strongly correlated with the SN host properties
\citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed
since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
the ``A+B'' model and the connection between SN stretch and host properties,
\cite{howell2007} first discussed the potential redshift drift of the SN stretch
distribution. In this paper we revisit this analysis with the most recent SNe~Ia
dataset.

@@

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf   & --          & 114 \\
        SDSS  & 0.20 (0.15) & 167 (82)\\
        PS1   & 0.30 (0.27) & 160 (122)\\
        SNLS  & 0.60 (0.55) & 102 (78)\\
        HST   & --          & 26 \\
        \hline
        Total & --          & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{Saunders2020}
\nn{\textit{Greg: for now you may need to cite SNEMO and SUGAR to remove this
``?''}}, see also Aldering in prep. and Ponder in prep.). The spectroscopic
follow-up was done over a redshift range of $0.02 < z < 0.09$~\citep[as
in][]{rigault2018}, while the search was much deeper. As such, these SNe~Ia are
assumed to be a random sampling of the underlying SN population. The SNfactory
sample is particularly useful for studying SN property drift, as it enables us
to have a large SN~Ia sample at $z<0.1$.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{multline}\label{eq:stretchz}
    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
\end{multline}

@@

Given the probability $y^i$ that a given SN is young, and assuming our Base (see
Section~\ref{sec:basemodel}), the probability to measure a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
\end{align}

@@

For the SNfactory sample, we can readily set $y^i = p_y^i$, the probability to
have $\log \mathrm{LsSFR} \geq -10.82$ (see Fig.~\ref{fig:stretchlssfr}), to
minimize Eq.~\ref{eq:likelihood} with respect to~$\vec{\theta}$. Results on
fitting the SNf SNe with this model are shown Table~\ref{tab:modelresults} and
illustrated in Fig.~\ref{fig:stretchlssfr}.

@@

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch (mean of Eq.~\ref{eq:stretchz})
illustrated as a blue band in Fig.~\ref{fig:modelall}. We see in this figure
that the measured mean SN~Ia stretch per redshift bins of equal sample size
closely follows our redshift drift modeling. This is indeed what is expected if
old environments favor low SN stretches \citep[e.g.][]{howell2007} \emph{and} if
the fraction of old SNe~Ia declines as a function of redshift. See
Section~\ref{sec:results} for a more quantitative discussion.

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift. Markers show the mean stretch measured in
        redshift bins of equal sample size, indicated in light gray at the
        bottom of each redshift bins. Full and light markers are used when
        considering the fiducial or the conservative samples, respectively. The
        orange horizontal line represents the mean stretch of the fiducial
        sample, illustrating the expectation if the SN stretch distribution is
        not evolving with redshift. Best fits of our Base drifting model are
        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
        the conservative one or the SNfactory dataset only, respectively; all
        are compatible. The light-blue band illustrates the amplitude of the
    error of the best fit model when considering the fiducial dataset.}
    \label{fig:modelall}
\end{figure*}

@@

Finally, for the sake of completeness, we also consider simple the ``Gaussian''
and ``Asymmetric'' Gaussian non-drifting models.

@@

Since the various models have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes extra degrees of
freedom to avoid over-fitting the data, and is defined as follow:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k
\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model is the one
with the smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
a Gaussian probability distribution.

We first note that the best model (with smallest AIC) is the so-called Base
model, implementing the \cite{rigault2018} description of the stretch
distribution and its redshift evolution. This is therefore the reference model,
both on the fiducial and conservative samples.

@@

Surprisingly, the sample-based Gaussian asymmetric modeling used by current
implementations of the BBC technique \citep{scolnic2016, kessler2017}, is one of
the worst modelisations of our analysis (see Section~\ref{sec:results}). While
its $-2\ln(L)$ is the smallest of all redshift-independent models (but still
$-11.5$ worse than the reference Base model), it is strongly penalized for
requiring 15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of
the analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted
as a probability $p=2\times 10^{-7}$ of being an as good representation of the
data as the Base model. We report in Table~\ref{tab:bbc} the samples' $\mu_0$
and $\sigma_{\pm}$ adjusted on the nominally selection-free samples using our
fiducial cuts (see Section~\ref{sec:sample}). We find our results in close
agreement with \cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a}
for PS1, who derived these model parameters using the full BBC formalism, using
numerous simulations to model the selection effects (see details e.g., Section~3
of \citealt{kessler2017}). The agreement between our fit of the asymmetric
Gaussians on the supposedly selection-free part of the samples and the results
derived using the BBC formalism supports our approach to get a sample with
negligible selection effects. If we were to use \cite{scolnic2016} and
\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
Base drifting model and the BBC modeling would go even deeper from $-32$ to
$-47$. We further discuss the consequence of this result for cosmology in
Section~\ref{sec:discussion}.

@@

Finally, \sout{considering the SNfactory dataset like any other (i.e.,} ignoring
the LsSFR measurements \nn{-- available only for the SNfactory dataset}, see
Section~\ref{sec:modeling} -- \sout{),} reduces the significance of the results
presented in this section, as expected. Yet, non-drifting models remain strongly
disfavored, and for instance, the best fitted sample-based Gaussian asymmetric
modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the data than
our Base drifting modeling.

@@

We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
underlying stretch distribution between the per-sample asymmetric modeling and
our Base drifting model for the PS1 sample. Our model is bimodal and the
relative amplitude of each mode depends on the redshift-dependent fraction of
old and young SNe~Ia in the sample: the higher the fraction of old SNe~Ia (at
lower redshift), the higher the amplitude of the old-specific low-stretch mode.
This redshift dependency is shown as blue to red underlying distributions in
Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
$x_1$ histogram follows our modeling defined as the sum of individual underlying
SN-redshift distributions. As expected, the two modeling approaches differ
mostly in the negative part of the SN stretch distribution. The asymmetric
Gaussian distribution goes through the middle of the bimodal distribution,
over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
redshifts. This means that the SN bias-corrected standardized magnitude
estimated at redshifts affected by selection effect would be biased by an
ill-modeling of the true underlying stretch distribution.

@@

The amplitude of this magnitude bias for cosmology is beyond the scope of this
paper given the complexity of the BBC analysis. It would require a full study
using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
asymmetric modeling as part of the BBC simulations.

Nonetheless, it is enlightening to study the expected mean stretch difference as
a function of redshift when considering that the underlying SN stretch
distribution follows either the per-sample Asymmetric Gaussian modeling or our
Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
we convert this difference in mean stretch $x_1$ to a difference in magnitude
using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
correspond to the actual impact on cosmological distances derived from SNe~Ia as
this simple stretch average comparison does not account for the amplitude of the
Malmquist bias correction to be applied as a function of redshift and surveys.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
    \caption{Difference of mean stretch standardization (in mag, using
        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
        Gaussian asymmetric modeling and our Base drifting model, as a function
        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
        considering the individual SN redshifts (no redshift cut applied). The
        orange markers represent their mean evolution in 30 regularly-spaced
        redshift bins. This must be multiplied by the amplitude of the Malmquist
        bias correction to derive the SN distance modulus bias affecting
        cosmology, see text. \nn{\textit{Apparently, there is some issue with
        the mean of all surveys combined. See Greg's comments.}}}
    \label{fig:magdrift}
\end{figure}

We see that the amplitude of the modeling difference is a few tens of millimag.
In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
stress that correctly modeling of potential SN redshift drift should be further
studied and care should be made when using samples affected by selection
effects.

@@

>>>>

\usepackage[dvipsnames]{xcolor}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{0.25, 0.50, 0}{#1}}}
\newcommand{\nnn}[1]{{\textcolor[rgb]{0.40, 0.0, 0.20}{#1}}}
\newcommand{\yc}[1]{{\textcolor{BrickRed}{#1}}}
\newcommand{\agk}[1]{\textcolor{purple}{AK: #1}}

\newcommand{\orcid}[1]{
\href{https://orcid.org/#1}{
\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}

@@

\author{
    N. Nicolas \thanks{n.nicolas@ipnl.in2p3.fr, equal contribution} \inst{1} 
    \and M. Rigault \thanks{m.rigault@ipnl.in2p3.fr, equal contribution} \inst{2}
    \orcid{0000-0002-8121-2560}
    \and\\ Y. Copin \inst{1}
    \and R. Graziani \inst{2}
    %HERE ALPHABETIC ORDER
    \and G. Aldering\inst{3}
    \and M. Briday\inst{1}
    \and J. Nordin\inst{4}
    \and A.~G. Kim\inst{3}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    % Gerard ne souhaite pas etre co-auteur.
}

@@

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    redshift drift. In this paper, we study the redshift evolution of SN~Ia
    \textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN
    property, to probe this drift. The SN stretch has been shown to strongly
    correlate with the SN's environment, notably with stellar age tracers. We
    model the underlying stretch distribution as a function of reshift, using
    the evolution of fraction of young and old SNe~Ia as predicted by
    \cite{rigault2018}, and assuming constant underlying stretch distribution
    for each age population made of Gaussian-mixtures.  We test our prediction
    against data from the literature chosen to have negligible selection effects
    to ensure that any observed drift is indeed astrophysical and not
    observational. We clearly demonstrate that the underlying SNe~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a much better description of the data than any
    non-drifting model, including the sample-based asymmetric distributions
    usually used to correct Malmquist bias. Our favored underlying stretch model
    is bimodal: a high-stretch mode shared by both young and old environments,
    and a low-stretch mode exclusive to old environments. The precise impact of
    the redshift evolution of the underlying SNe Ia population on cosmology
    remains to be studied. Yet, the astrophysical drift of the SN stretch
    distribution does affect current Malmquist bias corrections and thereby
    distances derived from SN affected by selection effects. We highlight that
    such a bias will increase with surveys covering larger redshift ranges,
    which is particularly important for LSST.}

@@

%\nn{We point that there has been sample-based approach meant to account for an
%unmodeled variation of host-related parameters thanks to multiple
%redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show
%aren't as effective as a direct modeling of such a drift}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). In \cite{rigault2018} we used the LsSFR to classify which are the
younger (those with a high LsSFR) and which are the older (those with low
LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN stretch has
been known to be strongly correlated with the SN host properties
\citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed
since \citep[e.g.][]{neill2009, sullivan2010, lampeitl2010, kelly2010,
gupta2011, dandrea2011, childress2013, rigault2013, pan2014, kim19}. Following
the ``A+B'' model and the connection between SN stretch and host properties,
\cite{howell2007} first discussed the potential redshift drift of the SN stretch
distribution. In this paper we revisit this analysis with the most recent SNe~Ia
dataset.

@@

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in parentheses.}
    \label{tab:sample}
    \begin{tabular}{l c c}
        \hline\hline
        Survey & $z_{lim}$ & $N_{\mathrm{SN}}$ \\
        \hline
        SNf   & --          & 114 \\
        SDSS  & 0.20 (0.15) & 167 (82)\\
        PS1   & 0.30 (0.27) & 160 (122)\\
        SNLS  & 0.60 (0.55) & 102 (78)\\
        HST   & --          & 26 \\
        \hline
        Total & --          & 569 (422) \\
        \hline
    \end{tabular}
\end{table}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. \mr{These 114 SNfactory SNe~Ia are thus part of the volume limited
part of the survey (Aldering in prep.).} As such, these SNe~Ia are assumed to be
a random sampling of the underlying SN population. The SNfactory sample is
particularly useful for studying SN property drift, as it enables us to have a
large SN~Ia sample at $z<0.1$.

@@

The stretch probability distribution function (pdf) of a given SN will be the
linear combination of the stretch distributions of these two population weighted
by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}).
But generally, the fraction of young SNe~Ia as a function of redshift is given
by $\delta(z)$ (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model
of the underlying distribution of SNe~Ia as a function of redshift $X_1(z)$ is
given by:
\begin{multline}\label{eq:stretchz}
    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
\end{multline}

@@

Given the probability $y^i$ that a given SN is young, and assuming our Base
\nn{model} (see Section~\ref{sec:basemodel}), the probability to measure a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg]
\end{align}

@@

\mr{Following \cite{rigault2018}, we set $p_y^i$ as one minus the cumulative
distribution function of the $\log(\mathrm{LsSFR})$ at
$-10.82\,\mathrm{dex}$. This is defining the probability that a SN projenitor
is classified as young (see further discussion about this assumption in Briday
et al. in~prep). Since the drift is negligible within the SNfactory dataset
($\Delta\delta(z)<3\%$ from the highest to the lowest redshift of the dataset),
one can readily use $y^i = p_y^i$ to minimize Eq.~\ref{eq:likelihood} with
respect to $\vec{\theta}$. Results on fitting the SNfactory SNe with this model
are shown Table~\ref{tab:modelresults} and illustrated in
Fig.~\ref{fig:stretchlssfr}.}

@@

%\agk{You don't define $p^i_y$. The language is is sloppy. What is relevant is
%the opposite, the probability that a galaxy with low mass gets the measured
%LsSFR. These are in general different. The solution you use is only correct
%for a model in which the rates for low and high-LsSFR galaxies is the same.
%The probability in general should include a prior on $l$.

%\begin{align}
%p(x^i_{o1},l_o^i | \mu_1,\mu_2,\sigma_1,\sigma_2,a) & = \int dl^i p(x^i_{o1},l_o^i | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
% & = \int dl^i p(x^i_{o1} | l^i, \mu_1,\mu_2,\sigma_1,\sigma_2,a) 
% p(l_o^i | li) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
%& = \int_{-\infty}^{-10.82} dl^i [a\mathcal{N}(\mu_1,\sigma_1{}^2) +
%    (1-a)\mathcal{N}(\mu_2,\sigma_2{}^2] 
% p(l_o^i | l^i) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber \\
%&+\int_{-10.82}^{\infty} dl^i\mathcal{N}(\mu_1,\sigma_1{}^2)
% p(l_o^i | l^i) 
%p(l^i| \mu_1,\mu_2,\sigma_1,\sigma_2,a ) \nonumber
%\end{align}
%Note that you use rates without LsSFR measurements, so you are doing something
%inconsistent here. There should be a $\delta{z}$ term here. However, since the
%rates are the same at $z=0.05$, maybe this doesn't matter so much.
%}

@@

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(resp. conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch \mr{(expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameters errors and their covariances}.
We see in this figure that the measured mean SN~Ia stretch per redshift bins of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.][]{howell2007} \emph{and} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf-hist_on.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift. Markers show the mean stretch measured in
        redshift bins of equal sample size, indicated in light gray at the
        bottom of each redshift bins. Full and light markers are used when
        considering the fiducial or the conservative samples, respectively. The
        orange horizontal line represents the mean stretch of the fiducial
        sample, illustrating the expectation if the SN stretch distribution is
        not evolving with redshift. Best fits of our Base drifting model are
        shown as blue, dashed-blue and gray, when fitted on the fiducial sample,
        the conservative one or the SNfactory dataset only, respectively; all
        are compatible. The light-blue band illustrates the amplitude of the
        error (incl. covariance) of the best fit model when considering the
        fiducial dataset. \mr{The redshift distributions of each sample are
        shown as light-histograms, see lower left legend.}}
    \label{fig:modelall}
\end{figure*}

@@

Finally, for the sake of completeness, we also consider Gaussian
and Asymmetric Gaussian non-drifting models.

@@

Since the various models have different degrees of freedom, we use the Akaike
Information Criterion \citep[AIC, e.g.][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes extra degrees of
freedom to avoid over-fitting the data, and is defined as follow:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k
\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model is the one
with the smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, the ones with
$5<\Delta\mathrm{AIC}<20$ are unfavored, and those with $\Delta\mathrm{AIC}>20$
are deemed excluded. This roughly corresponds to 2, 3 and 5~$\sigma$ limits for
a Gaussian probability distribution. \mr{However, we highlight that the best of
the compared models might not be the true underlying model.}

The best model (with smallest AIC) is the so-called Base model and thus is out
the reference model. This is true both on the fiducial and conservative samples.

@@

Surprisingly, the sample-based Gaussian asymmetric modeling used by current
implementations of the BBC technique \citep{scolnic2016, kessler2017} has one of
the highest AIC value in our analysis (see Section~\ref{sec:results}). While its
$-2\ln(L)$ is the smallest of all redshift-independent models (but still $-11.5$
worse than the reference Base model), it is strongly penalized for requiring
15~free parameters ($\mu_0, \sigma_{\pm}$~for each of the 5~samples of the
analysis). Hence, its $\Delta\mathrm{AIC}<-20$, which could be interpreted as a
probability $p=2\times 10^{-7}$ of being an as good representation of the data
as the Base model.
\mr{Remark that the Bayesian Information Criterion ($\mathrm{BIC} = -2\ln(L) +
k\ln(n)$, with $n$ is the number of datapoint) might be better suited than AIC
when comparin models to the sample-based one since it explicitely accounts for
the fact that each sample is fitted separately; the sample-based model BIC is
the sum of the BIC per sample. Doing so we find $\Delta\mathrm{BIC}=-48$, also
excluding the sample-base asymmetric Gaussian model as an as good description of
the data then the Base-model.}

%\agk{I think it does worse not because of the more parameters but that it has 5
%complexity penalties whereas the base model only has 1.  AIC doesn't account
%for the fact that each sample is fit independently.  It seems a fairer metric
%would be the BIC, where you adjust $n$ for each sample.  This would recover
%$\sim 16$ in the IC for this model.}

We report in Table~\ref{tab:bbc} the samples' $\mu_0$ and $\sigma_{\pm}$
adjusted on the nominally selection-free samples using our fiducial cuts (see
Section~\ref{sec:sample}). We find our results in close agreement with
\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
derived these model parameters using the full BBC formalism, using numerous
simulations to model the selection effects (see details e.g., Section~3 of
\citealt{kessler2017}). The agreement between our fit of the asymmetric
Gaussians on the supposedly selection-free part of the samples and the results
derived using the BBC formalism supports our approach to get a sample with
negligible selection effects. If we were to use \cite{scolnic2016} and
\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
Base drifting model and the BBC modeling would go even deeper from $-32$ to
$-47$. We further discuss the consequence of this result for cosmology in
Section~\ref{sec:discussion}.

@@

Finally, ignoring the LsSFR measurements --~available only for the SNfactory
dataset, see Section~\ref{sec:modeling}~-- reduces the significance of the
results presented in this section, as expected. Yet, non-drifting models remain
strongly disfavored, and for instance, the best fitted sample-based Gaussian
asymmetric modeling still is $\Delta\mathrm{AIC}<-10$ less representative of the
data than our Base drifting modeling.

@@

We illustrate in Fig.~\ref{fig:bbc_pdf_ps1} the prediction difference in the
underlying stretch distribution between the per-sample asymmetric modeling and
our Base drifting model for the PS1 sample. Our model is bimodal and the
relative amplitude of each mode depends on the redshift-dependent fraction of
old and young SNe~Ia in the sample: the higher the fraction of old SNe~Ia (at
lower redshift), the higher the amplitude of the old-specific low-stretch mode.
This redshift dependency is shown as blue to red underlying distributions in
Fig.~\ref{fig:bbc_pdf_ps1} for redshift ranges covered by PS1. The observed
$x_1$ histogram follows our modeling defined as the sum of individual underlying
SN-redshift distributions. As expected, the two modeling approaches differ
mostly in the negative part of the SN stretch distribution. The asymmetric
Gaussian distribution goes through the middle of the bimodal distribution,
over-estimating the number of SNe~Ia at $x_1\sim-0.7$ and under-estimating it at
$x_1\sim-1.7$ in comparison to our Base drifting model for typical PS1 SN
redshifts. This means that the SN bias-corrected standardized magnitude
estimated at a redshift affected by selection effect would be biased by an
ill-modeling of the true underlying stretch distribution.

@@

The amplitude of this magnitude bias for cosmology is beyond the scope of this
paper given the complexity of the BBC analysis. It would require a full study
using our Base model (Eq.~\ref{eq:stretchz}) in place of the sample-based
asymmetric modeling as part of the BBC simulations.
\mr{However, we already highlight that even if a non-drifting sample-based model
could provide comparable result in the volume-limited part of the various
samples, these models would differ when extrapolating at higher redshift,
precisely where the  underlying will matter for correcting Malmquist biases.}

%Nonetheless, it is enlightening to study the expected mean stretch difference
%as a function of redshift when considering that the underlying SN stretch
%distribution follows either the per-sample Asymmetric Gaussian modeling or our
%Base drifting model. This is shown Fig.~\ref{fig:magdrift}, where, for clarity
%we convert this difference in mean stretch $x_1$ to a difference in magnitude
%using $\alpha=0.156$ \citep{scolnic2018a}. We stress here that this does not
%correspond to the actual impact on cosmological distances derived from SNe~Ia
%as this simple stretch average comparison does not account for the amplitude of
%the Malmquist bias correction to be applied as a function of redshift and
%surveys.

%\agk{ The AG modeling contains Malmquist bias. To be fair you should plot Base
%with the survey-specific Malmquist bias. There is something strange going on at
%$z<0.2$, can you add SNfactory to this plot? I think you fill find that the
%systematic difference is driven by differences in the $x_1$ distribution of
%SNf.}

%\begin{figure}
%    \centering
%    \includegraphics[width=\linewidth]{Article_figures/BBC_-stretchevol_grouped.pdf}
%    \caption{Difference of mean stretch standardization (in mag, using
%        $\alpha=0.156$ from \citealt{scolnic2018a}) between the per-sample
%        Gaussian asymmetric modeling and our Base drifting model, as a function
%        of redshift, considering the SDSS, PS1 and SNLS datasets. The colored
%        markers show the difference for SDSS, PS1 and SNLS samples (see legend)
%        considering the individual SN redshifts (no redshift cut applied). The
%        orange markers represent their mean evolution in 30 regularly-spaced
%        redshift bins. This must be multiplied by the amplitude of the Malmquist
%        bias correction to derive the SN distance modulus bias affecting
%        cosmology, see text. \nn{\textit{Apparently, there is some issue with
%        the mean of all surveys combined. See Greg's comments.}}}
%    \label{fig:magdrift}
%\end{figure}

%We see that the amplitude of the modeling difference is a few tens of millimag.
In the era of modern cosmology, where we aim at probing $w_0$ at a sub-percent
level and $w_a$ at the ten-percent precision \citep[e.g.,][]{lsstpaper}, we
stress that correctly modeling of potential SN redshift drift should be further
studied and care should be made when using samples affected by selection
effects.

@@

\bibitem[Aldering et al.(2020)]{aldering2020} Aldering, G., Antilogus, P.,
Aragon, C., et al.\ 2020, Research Notes of the American Astronomical Society,
4, 63
